Effect of quality-adjusted production function inputs

### 2015-04-29 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Improved authors. Improved text.

\author[Leeds]{Paul Brockway}
\author[IST]{Tiago Domingos}
\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}

% \address[IST]{IST, Lisbon, Portugal}
% \address[Leeds]{Leeds University, Leeds, UK}
% \cortext[cor1]{Corresponding author}
% \address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}

@@

Note that in the fitting process for the Cobb-Douglas equation,
$\texttt{scale} = \theta$, 
$\alpha_1 = \alpha$,
$\alpha_2 = \beta$, and
$\alpha_3 = \gamma$,
when the \texttt{formula} is specified as 
%
\begin{equation} \label{eq:formula_kle}
  \texttt{y} \sim \texttt{k + l + e + time}.
\end{equation}
%
$\theta$ is expected to be near unity.

Note that for the CES model, $\texttt{scale} = \gamma$ and is expected to be close to unity. 
Equations for $\alpha_1$, $\alpha_2$, and $\alpha_3$ are given in Table \ref{tab:CES_abg_equations}.

\begin{table} \caption{Equations for $\alpha$, $\beta$, and $\gamma$ for the various CES nestings, 
                        provided that the \texttt{formula} is specified as \texttt{y~$\sim$~k~+~l~+~e~+~time}.} 
\label{tab:CES_abg_equations} 
  \begin{center}
    \begin{tabular}{c c c c c} 
      \toprule
      Nesting         & \texttt{nest}     & $\alpha_1$ ($\alpha$)   & $\alpha_2$ ($\beta$)    & $\alpha_3$ ($\gamma$)       \\
      \midrule
      ($kl$) + ()     & \texttt{c(1,2)}   & $\delta_1$              & $1 - \delta_1$          & 0          \\
      ($kl$) + ($e$)  & \texttt{c(1,2,3)} & $\delta \delta_1$       & $\delta (1 - \delta_1)$ & $1 - \delta$ \\   
      ($le$) + ($k$)  & \texttt{c(2,3,1)} & $1 - \delta$            & $\delta \delta_1$       & $\delta (1 - \delta_1)$ \\
      ($ek$) + ($l$)  & \texttt{c(3,1,2)} & $\delta (1 - \delta_1)$ & $1 - \delta$            & $\delta \delta_1$ \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

Table~\ref{tab:lambda_CD_PT} shows
the effect of both quality-corrected factors of production and energy.
We find that both including energy and quality-adjusting the factors of production 
reduce the Solow residual (\texttt{lambda}). 
Quality-adjustment gets part of the way toward eliminating the Solow residual.
Including energy (whether or not the factors of production are quality-adjusted)
reduces the Solow residual significantly.

@@

Note that the contribution of capital (\texttt{alpha\_1})
always shifts to energy (\texttt{alpha\_3}) when energy is included 
as a factor of production.

@@

<<UK C-D>>=
@

@@

For the UK, the quality-adjusted, with energy Solow residual is 
\Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda * 100}%,
a very small number indeed.
For the UK, it appears that both quality-adjustment and addition of energy as a factor of production
can effectively eliminate the Solow residual,
reducing it to \Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda}.

Interestingly, the Solow residual acutally increases from 
\Sexpr{naturalCoef(models$UK$unadj$noE$CD)$lambda} to
\Sexpr{naturalCoef(models$UK$unadj$withE$CD)$lambda}
for unadjusted factors of production when energy is added.
To find out why, let's look at all the coefficients.

@@

First, look at the CES model without energy.
Nesting is unabmiguous, so this fit does not need to be repeated
for each nesting option.

@@

First, look at the CES model without energy.
Nesting is unabmiguous, so this fit does not need to be repeated
for each nesting option.

>>>>

\author[IST]{João dos Santos}
\author[Leeds]{Paul Brockway}
\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}
\author[Leeds]{Marco Sekai}
\author[IST]{Tiago Domingos}
\author[Leeds]{Julia Steinberger}

\address[IST]{IST, Lisbon, Portugal}
\address[Leeds]{Leeds University, Leeds, UK}
\address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}
\cortext[cor1]{Corresponding author}

@@

Note that in the fitting process for the Cobb-Douglas equation,
$\texttt{scale} = \theta$, 
$\alpha_1 = \alpha$,
$\alpha_2 = \beta$, and
$\alpha_3 = \gamma$,
when the \texttt{formula} is specified as 
%
\begin{equation} \label{eq:formula_kle}
  \texttt{y} \sim \texttt{k + l + e + time}.
\end{equation}
%
Note that for the CES model, $\texttt{scale} = \gamma$ and is expected to be close to unity.

For the CES model, values of $\alpha_1$, $\alpha_2$, and $\alpha_3$ are calculated by
$\alpha_1 = \delta_1$, $\alpha_2 = 1 - \delta_1$, and $\alpha_3 = 0$ 
for the CES model with two factors of production
and by 
$\alpha_1 = \delta \delta_1$, $\alpha_2 = \delta(1 - \delta_1)$, and $\alpha_3 = 1 - \delta$
for the CES model with three factors of production.
The values of $\alpha_1$, $\alpha_2$, and $\alpha_3$ are interpreted as
$\alpha$, $\beta$, and $\gamma$ (coefficients for capital, labor, and energy, respectively)
as shown in Table \ref{tab:CES_abg_equations}, 
assuming factors of production are specified in the fitting \texttt{formula} as
\texttt{y~$\sim$~k~+~l~+~e~+~time}.

\begin{table} \caption{Equations for $\alpha$, $\beta$, and $\gamma$ 
                        (importance of capital, labor, and energy, respectively)
                        for the various CES nestings, 
                        provided that the \texttt{formula} is specified as \texttt{y~$\sim$~k~+~l~+~e~+~time}.} 
\label{tab:CES_abg_equations} 
  \begin{center}
    \begin{tabular}{c c c c c} 
      \toprule
      Nesting         & \texttt{nest}     & $\alpha$                & $\beta$                 & $\gamma$       \\
      \midrule
      ($kl$) + ()     & \texttt{c(1,2)}   & $\delta_1$              & $1 - \delta_1$          & 0          \\
      ($kl$) + ($e$)  & \texttt{c(1,2,3)} & $\delta \delta_1$       & $\delta (1 - \delta_1)$ & $1 - \delta$ \\   
      ($le$) + ($k$)  & \texttt{c(2,3,1)} & $1 - \delta$            & $\delta \delta_1$       & $\delta (1 - \delta_1)$ \\
      ($ek$) + ($l$)  & \texttt{c(3,1,2)} & $\delta (1 - \delta_1)$ & $1 - \delta$            & $\delta \delta_1$ \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

Table~\ref{tab:lambda_CD_PT} shows
the effect of both quality-corrected factors of production and energy.
We find that including energy as a factor of production 
reduces the absolute value of the Solow residual (\texttt{lambda})
whether or not the factors of production are quality-adjusted. 
Quality-adjustment gets part of the way toward eliminating the Solow residual.

@@

Note that for Portugal the contribution of capital (\texttt{alpha\_1})
always shifts to energy (\texttt{alpha\_3}) when energy is included 
as a factor of production.

@@

Table~\ref{tab:lambda_CD_UK} shows fitted Cobb-Douglas coefficients for the UK.

@@

For the UK, the quality-adjusted, with energy Solow residual is 
\Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda * 100}\%,
a very small number indeed.
For the UK, it appears that \emph{both} quality-adjustment and addition of energy as a factor of production
are required to effectively eliminate the Solow residual,
reducing it to \Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda}.

Adding energy alone actually \emph{in}creases the Solow residual from 
\Sexpr{naturalCoef(models$UK$unadj$noE$CD)$lambda} to
\Sexpr{naturalCoef(models$UK$unadj$withE$CD)$lambda}.
The details for each model with unadjusted factors of production are shown below.

@@

First, look at the CES model without energy.
Nesting is unabmiguous, so this fit does not need to be repeated
for each nesting option. 
(This means that the top rows of Tables~\ref{tab:lambda_CES-kle_PT}--\ref{tab:lambda_CES-ekl_PT}
are identical.)

@@

First, look at the CES model without energy.
Nesting is unabmiguous, so this fit does not need to be repeated
for each nesting option. 
(This means that the top rows of Tables~\ref{tab:lambda_CES-kle_UK}--\ref{tab:lambda_CES-ekl_UK}
are identical.)
